---
title: "North Temperate Lakes LTER"
author: "Desik Somasundaram"
date: "8/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
```

## Read in data

```{r}
ice_data <- read_csv(here("data", "ntl33_v7.csv" ), na = "-999")
met_data <- read_csv(here("data", "ntl20_v6.csv" ))

```
## Ice Duration visualization

```{r}
ggplot(data = ice_data, aes(x = lakeid, y = ice_duration)) +
  geom_violin(fill = NA) +
  geom_jitter(aes(color = lakeid), width = 0.2) 
  
```
## Find mean ice cover duration

```{r}
ice_duration_yearly <- ice_data %>%
  rename(year= year4) %>% 
  group_by(year) %>% 
  summarize(mean_ice_duration = mean(ice_duration , na.rm = TRUE))
  
```

## Exploring Mean Ice Cover Duration

```{r}
ggplot(data = ice_duration_yearly, aes(x = year, y = mean_ice_duration)) +
  geom_point() +
  geom_line()
  
```
The mean ice duration indicates that the duration by year is generally cyclic with a general decreasing trend. 

## Mean Air Temperature 

```{r}
mean_winter <- met_data %>% 
  filter(month %in% c(12, 1, 2)) %>% 
  rename(year= year4) %>% 
  group_by(year) %>% 
  summarize(mean_airtemp_c = mean(ave_air_temp_adjusted, na.rm = TRUE))

```

## Exploring Mean Air Temp

```{r}
ggplot(data = mean_winter, aes(x = year, y = mean_airtemp_c)) +
  geom_point() +
  geom_line()
  
```


The mean air temperature indicates that the air temperature by year generally fluctuates drastically with a general increasing trend. 

## Joining air temperature and ice duration

```{r}
met_ice_mean <- left_join(mean_winter, ice_duration_yearly)
```
## Exploring Mean Air Temp and Duration Relationship

```{r}
ggplot(data = met_ice_mean, aes(x = mean_airtemp_c, y =  mean_ice_duration)) +
  geom_point() 

  
```

There is a generally decreasing trend as the mean air temperature increases, the mean ice duration decrease. However, the correlation is weak. 


## Exploring Thaw Dates


```{r}

post1970_ice_data <- ice_data %>%
  filter(year4 >= 1970) %>%
  mutate(ice_off = lubridate::ymd(ice_off)) %>% 
  mutate(yearday = lubridate::yday(ice_off))


ggplot(data = post1970_ice_data, aes(x = year4, y = yearday)) +
  geom_point() +
  geom_line()

```

There is no significant observable trend in the ice off day of the year as the years increase. 
